\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2017-05-17 redaction/dhp_sept2016.tex
Discussion sur le choix du début
1974 ou 1980 ? Cela ne change pas grand chose.

In this section, we investigate the role of the additive component in total transport costs in a historical perspective. As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, which displays the respective shares of additive and iceberg components in transport costs over time (by transport mode).
Let us consider first the trend in total transport costs (ie, considering the upper line in Figure \ref{fig:decomp_TC_3d}). Two main comments can be made. First, both air and ocean shipping exhibit a downward trend in overall transport costs since 1974. \todo{Guillaume, ameliorer les choses} If we start the analysis from 1980, the decrease is roughly 1.5-2 percentage points for air transport costs, and 2-3 percentage points for vessel transport costs (depending whether we consider mean or median estimates). These results are consistent with \citet{hummels2007}.

>>>>

In this section, we investigate the role of the additive component in total transport costs in a historical perspective. As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, which displays the respective shares of additive and iceberg components in transport costs over time (by transport mode).
Let us consider first the trend in total transport costs (ie, considering the upper line in Figure \ref{fig:decomp_TC_3d}). Both air and ocean shipping exhibit a downward trend in overall transport costs since 1974. (-2.1\% a year for mean air transport costs and -2.0\% a year for mean ocean transport costs).
\footnote{One may worry that this result springs from high oil-shock related transport costs in 1974. However, computing the time trends from 1980 does not change dramatically the picture. The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean ocean transport costs. In order to go as far as possible into the past, this paper will focus on the results from 1974}
 \todo{GD : j'ai amélioré. Qu'en penses-tu ?}
These results are consistent with \citet{hummels2007}.
